---
title: "S&P Momentum and Contrarian"
output: html_notebook
---
S&P 500 Contrarian and Momentum Strategies 2010-2018

```{r}
library(zoo)
library(dplyr)
```

```{r}
sp2010<- read.csv("/Users/nastyademina/Desktop/r studio/prices_2010-2016.csv")
sp2013 <- read.csv("/Users/nastyademina/Desktop/r studio/all_stocks_5yr.csv")
```

```{r}
#making sure we have the same tickers in two data sets
test<-sp2010[which(sp2010$symbol %in% unique(sp2013$Name)),]
sp2013 <- sp2013[which(sp2013$Name %in% unique(test$symbol)),]

#changing the dates
library("lubridate")
test$date <- gsub(x=test$date,pattern="00:00:00",replacement="",fixed=T)
test$date <- as.Date(as.character(test$date), format="%Y-%m-%d")
sp2013$date <- as.Date(as.character(sp2013$date), format="%Y-%m-%d")
```

```{r}
#merge two data frames
sp2013 <- sp2013[which(sp2013$date>'2016-12-31'),]
colnames(sp2013) <- c('date','open', 'high', 'low','close', 'volume', 'symbol')
sp2013 <- sp2013[,c('date','symbol','open','close','low','high','volume')]
sp <- rbind(test, sp2013)

#subsetting the data (we only will look at the close price, so we need date, closing price and ticker)
sp500 <- subset(sp, select=c('date','close','symbol'))

#reshape the data to get the shape of the data frame that we need
sp500 <- reshape(sp500, timevar = "symbol", idvar = "date", direction = "wide")
```

```{r}
#for simplicity get rid of columns with NA (stocks that have some missing observations)
list_na <- c(0)
for (i in 2:ncol(sp500)){
  list_na[i] <- sum(is.na(sp500[,i]))
}
c <- list_na==0

#new dataframe contains only 438 stocks that stayed in S&P from 2010 to 2018
sp500 <- sp500[,c]
sp500 <- sp500[order(sp500$date),]
```

Contrarian Strategy
```{r}
#creating a data frame of returns
df<-as.data.frame(subset(sp500, select = "date"))

for (i in 2:ncol(sp500)){
  ret <- (sp500[,i] - lag(sp500[,i]))/lag(sp500[,i])
  df <- cbind(df,ret)
}

df <- df[-1,] # cut first row of NA
dates <- as.data.frame(df$date[2:nrow(df)])
colnames(dates)[1] <- "date"
```

```{r,warning=FALSE}
#calculating market return every day
df$mean <- apply(df[,-1], 1, FUN=mean)
market <- as.data.frame(df$mean[2:nrow(df)])
colnames(market)[1] <- "Market Return"
```

```{r}
#defining signal
contr <- -(df[,c(2:(ncol(df)-1))]-df$mean)

#function to sell winners and buy losers
contrarian<- function(x){
  index <- order(x, decreasing=TRUE)
  x.sort <- sort(x, decreasing=TRUE)
  constant <- ceiling(length(x)/10)
  vec <- c(rep.int(1/constant, constant), rep.int(0,length(x)-2*constant),rep.int(-1/constant, constant))
  x.sort <- rbind(x.sort,vec)
  x.sort <- x.sort[-1,]
  x[index] <- x.sort
  return(x)
}
#calculating weights
weight_contr <- apply(contr, 1, contrarian)
weight_contr <- t(weight_contr)

#cut the first row of returns and the last row of weights since we are using lagged weights to calculate our returns 
return <- df[-1, 2:(ncol(df)-1)]
weight_contr <- weight_contr[1:(nrow(weight_contr)-1),]
port_return <- return*weight_contr

sum(port_return[1,]!=0)
sum(port_return[19,]!=0)

#calculating the strategy profit
port_return.contr <- apply(port_return, 1, sum)
cum.ret.contr <- cumprod(1+port_return.contr)*100
```

Measuring Performance
```{r}
ann_return.contr <- mean(port_return.contr)*252*100
ann_return.contr
std_return.contr <- sd(port_return.contr)*sqrt(252)*100
std_return.contr
sharpe.contr <- ann_return.contr/std_return.contr
sharpe.contr
```

#Graphs
##Stocks mean and cumulative mean returns
```{r}
library(ggplot2)
colnames(df)<-c(colnames(sp500),"mean")
mean_ret<-subset(df, select = c("date","mean"))
mean_ret<-mutate(mean_ret, cum_ret = cumprod(1+mean))
ggplot(mean_ret)+geom_line(aes(x=date,y=mean))+theme_minimal()+xlab("Date")+ylab("Returns") + ggtitle("Average Returns of 438 Stocks in S&P500")

ggplot(mean_ret)+geom_line(aes(x=date,y=cum_ret))+theme_minimal() + xlab("Date")+ylab("Returns") + ggtitle("Cumulative Average Return of 438 Stocks in S&P500")
```

## Autocorrelation of stocks
```{r}
acf_fun<-function(x){
  acf(x, plot = FALSE, lag.max = 1)[1]
}

acf_sp<-c(0)
for(i in 2:(ncol(df)-1)){
  acf_sp[i]<-as.numeric(unlist(acf_fun(df[,i]))[1])
}
mean(acf_sp)

hist(acf_sp, main = "Histogram of 1 lag autocorrelations of stocks in S&P", xlab = "Autocorrelation")
```

## Portfolio Returns
```{r}
dates<-as.data.frame(df$date[2:nrow(df)])
port.returns<-cbind(dates,port_return.contr,cum.ret.contr)
colnames(port.returns)<-c("date","port_ret","cum_ret")
ggplot(port.returns) + geom_line(aes(x =date,y=port_ret ))+theme_minimal()+xlab("Date") + ylab("Returns") + ggtitle("Daily Portfolio Returns of Contrarian Strategy") 

ggplot(port.returns) + geom_line(aes(x =date,y=cum_ret ))+theme_minimal()+xlab("Date") + ylab("Returns") + ggtitle("Cumulative Portfolio Returns of Contrarian Strategy") 
```

Momentum strategy for 6 months
```{r}
#creating a data frame of returns
mom<-as.data.frame(subset(sp500, select = "date"))

for (i in 2:ncol(sp500)){
  ret <- (sp500[,i] - lag(sp500[,i]))/lag(sp500[,i])
  mom <- cbind(mom,ret)
}

mom <- mom[-1,-1] # cut first row of NA
test <- mom+1
test <- rollapply(test,126,FUN=prod, by=126) #calculate cumulative returns every six months 
#we look at the cumulative return for the past six months and based on them calculate the signal 
#we use that signal for the next six months
test <- as.data.frame(test)
test$mean <- apply(test, 1, FUN=mean)
test <- test[1:(nrow(test)-1),] 

mat <- as.data.frame(matrix(NA,nrow(test)*126,ncol(test)))
for (i in 0:(nrow(test)-1)){
  mat[((126*i+1):(126+126*i)),] <- test[(i+1),]
}
```

```{r}
#signal
colnames(mat)[ncol(mat)] <- "mean"
m <- (mat[,c(1:(ncol(mat)-1))]-mat$mean)

#function to sell winners and buy losers
momentum<- function(x){
  index <- order(x, decreasing=TRUE)
  x.sort <- sort(x, decreasing=TRUE)
  constant <- ceiling(length(x)/10)
  vec <- c(rep.int(1/constant, constant), rep.int(0,length(x)-2*constant),rep.int(-1/constant, constant))
  x.sort <- rbind(x.sort,vec)
  x.sort <- x.sort[-1,]
  x[index] <- x.sort
  return(x)
}

weight_mom <- apply(m, 1, momentum)
weight_mom <- t(weight_mom)

return <- mom[-(1:126), 1:(ncol(mom))]
return <- return[1:(nrow(return)-22), 1:(ncol(return))]
weight_mom <- weight_mom[1:(nrow(weight_mom)),]
port_return.mom <- return*weight_mom

sum(port_return.mom[125,]!=0)
sum(port_return.mom[19,]!=0)

#calculating the strategy profit
port_return.mom <- apply(port_return.mom, 1, sum)
cum.ret.mom <- cumprod(1+port_return.mom)*100
```

Measuring Performance
```{r}
ann_return.mom <- mean(port_return.mom)*252*100
ann_return.mom
std_return.mom <- sd(port_return.mom)*sqrt(252)*100
std_return.mom
sharpe.mom <- ann_return.mom/std_return.mom
sharpe.mom
```
#Graphs
## Portfolio Returns
```{r}
dates<-as.data.frame(subset(sp500[2:nrow(sp500),], select = "date"))
dates<-as.data.frame(dates[-(1:126),])
dates<-as.data.frame(dates[1:(nrow(dates)-22),])

port.returns<-cbind(dates,port_return.mom,cum.ret.mom)
colnames(port.returns)<-c("date","port_ret","cum_ret")
ggplot(port.returns) + geom_line(aes(x =date,y=port_ret ))+theme_minimal()+xlab("Date") + ylab("Returns") + ggtitle("Daily Portfolio Returns of Momentum Strategy") 

ggplot(port.returns) + geom_line(aes(x =date,y=cum_ret ))+theme_minimal()+xlab("Date") + ylab("Returns") + ggtitle("Cumulative Portfolio Returns of Momentum Strategy") 
```



Momentum strategy for 6 months and rebalance every 3 months
```{r}
#creating a data frame of returns
mom<-as.data.frame(subset(sp500, select = "date"))

for (i in 2:ncol(sp500)){
  ret <- (sp500[,i] - lag(sp500[,i]))/lag(sp500[,i])
  mom <- cbind(mom,ret)
}

mom <- mom[-1,-1] # cut first row of NA
test <- mom+1
test <- rollapply(test,126,FUN=prod, by=63) #calculate cumulative returns every six months 
#we look at the cumulative return for the past six months and based on them calculate the signal 
#we use that signal for the next three months
test <- as.data.frame(test)
test$mean <- apply(test, 1, FUN=mean)
test <- test[1:(nrow(test)-1),] 

mat <- as.data.frame(matrix(NA,nrow(test)*63,ncol(test)))
for (i in 0:(nrow(test)-1)){
  mat[((63*i+1):(63+63*i)),] <- test[(i+1),]
}
```

```{r}
colnames(mat)[ncol(mat)] <- "mean"
m <- (mat[,c(1:(ncol(mat)-1))]-mat$mean)

#function to sell winners and buy losers
momentum<- function(x){
  index <- order(x, decreasing=TRUE)
  x.sort <- sort(x, decreasing=TRUE)
  constant <- ceiling(length(x)/10)
  vec <- c(rep.int(1/constant, constant), rep.int(0,length(x)-2*constant),rep.int(-1/constant, constant))
  x.sort <- rbind(x.sort,vec)
  x.sort <- x.sort[-1,]
  x[index] <- x.sort
  return(x)
}

weight_mom <- apply(m, 1, momentum)
weight_mom <- t(weight_mom)

return <- mom[-(1:126), 1:(ncol(mom))]
return <- return[1:(nrow(return)-22), 1:(ncol(return))]
weight_mom <- weight_mom[1:(nrow(weight_mom)),]
port_return.mom <- return*weight_mom

sum(port_return.mom[125,]!=0)
sum(port_return.mom[19,]!=0)

#calculating the strategy profit
port_return.mom <- apply(port_return.mom, 1, sum)
```

Measuring Performance
```{r}
ann_return.mom <- mean(port_return.mom)*252*100
ann_return.mom
std_return.mom <- sd(port_return.mom)*sqrt(252)*100
std_return.mom
sharpe.mom <- ann_return.mom/std_return.mom
sharpe.mom
```

Transaction costs analysis
Contrarian
```{r}
df<-as.data.frame(subset(sp500, select = "date"))

for (i in 2:ncol(sp500)){
  ret <- (sp500[,i] - lag(sp500[,i]))/lag(sp500[,i])
  df <- cbind(df,ret)
}

df <- df[-1,] # cut first row of NA
dates <- as.data.frame(df$date[2:nrow(df)])
colnames(dates)[1] <- "date"
```

```{r,warning=FALSE}
#calculating mean
df$mean <- apply(df[,-1], 1, FUN=mean)
market <- as.data.frame(df$mean[2:nrow(df)])
colnames(market)[1] <- "Market Return"
```

```{r}
contr <- -(df[,c(2:(ncol(df)-1))]-df$mean)

#function to sell winners and buy losers
contrarian<- function(x){
  index <- order(x, decreasing=TRUE)
  x.sort <- sort(x, decreasing=TRUE)
  constant <- ceiling(length(x)/10)
  vec <- c(rep.int(1/constant, constant), rep.int(0,length(x)-2*constant),rep.int(-1/constant, constant))
  x.sort <- rbind(x.sort,vec)
  x.sort <- x.sort[-1,]
  x[index] <- x.sort
  return(x)
}

weight_contr <- apply(contr, 1, contrarian)
weight_contr <- t(weight_contr)
cum.ret.contr.1 <- c(100,cum.ret.contr)
```

```{r}
#calculating transaction costs
trans <- cum.ret.contr.1*weight_contr
trans <- rbind(0, trans)

transaction<-data.frame(NA)

for (i in 1:ncol(trans)){
  cost <- abs(trans[,i] - lag(trans[,i]))
  transaction<- cbind(transaction,cost)
}

transaction <- transaction[-1,-1]
trans_cost <- 0.01*apply(transaction, 1, sum)

trans_cost <- c(0,trans_cost[1:(length(trans_cost)-1)] )
#####################1
#cum.ret.contr1. <- cum.ret.contr.1[1:(length(cum.ret.contr.1)-1)]
#net_profit.contr <- net_profit.contr[1:(length(net_profit.contr)-1)]
net_profit.contr <- cum.ret.contr.1-trans_cost

net_return.contr<-(net_profit.contr/lag(net_profit.contr))-1

plot(net_return.contr, type = "l")
```

Net profit graph
```{r}
plot(net_profit.contr, type='l')
```
Perfomance
```{r}
ann_return.contr <- mean(net_return.contr, na.rm = T)*252*100
ann_return.contr
std_return.contr <- sd(net_return.contr, na.rm = T)*sqrt(252)*100
std_return.contr
sharpe.contr <- ann_return.contr/std_return.contr
sharpe.contr
```
[1] -8.516652
[1] 15.98407
[1] -0.5328214

Transaction costs
Momentum
```{r}
mom<-as.data.frame(subset(sp500, select = "date"))

for (i in 2:ncol(sp500)){
  ret <- (sp500[,i] - lag(sp500[,i]))/lag(sp500[,i])
  mom <- cbind(mom,ret)
}

mom <- mom[-1,-1] # cut first row of NA
test <- mom+1
test <- rollapply(test,126,FUN=prod, by=126) #calculate cumulative returns every six months 
#we look at the cumulative return for the past six months and based on them calculate the signal 
#we use that signal for the next six months
test <- as.data.frame(test)
test$mean <- apply(test, 1, FUN=mean)
test <- test[1:(nrow(test)-1),] 

mat <- as.data.frame(matrix(NA,nrow(test)*126,ncol(test)))
for (i in 0:(nrow(test)-1)){
  mat[((126*i+1):(126+126*i)),] <- test[(i+1),]
}
```

```{r}
#signal
colnames(mat)[ncol(mat)] <- "mean"
m <- (mat[,c(1:(ncol(mat)-1))]-mat$mean)

#function to sell winners and buy losers
momentum<- function(x){
  index <- order(x, decreasing=TRUE)
  x.sort <- sort(x, decreasing=TRUE)
  constant <- ceiling(length(x)/10)
  vec <- c(rep.int(1/constant, constant), rep.int(0,length(x)-2*constant),rep.int(-1/constant, constant))
  x.sort <- rbind(x.sort,vec)
  x.sort <- x.sort[-1,]
  x[index] <- x.sort
  return(x)
}

weight_mom <- apply(m, 1, momentum)
weight_mom <- t(weight_mom)


#calculating the strategy profit
cum.ret.mom.1 <- c(100,cum.ret.mom)
cum.ret.mom.1 <- cum.ret.mom.1[1:(length(cum.ret.mom.1)-1)]
```

```{r}
#calculating transaction costs
trans <- cum.ret.mom.1*weight_mom

transaction<-data.frame(NA)

for (i in 1:ncol(trans)){
  cost <- abs(trans[,i] - lag(trans[,i]))
  transaction<- cbind(transaction,cost)
}

transaction <- transaction[-1,-1]
trans_cost.mom <- 0.01*apply(transaction, 1, sum)

trans_cost.mom <- c(0,trans_cost.mom[1:(length(trans_cost.mom))])

net_profit.mom <- cum.ret.mom.1-trans_cost.mom
net_return.mom<-net_profit.mom/lag(net_profit.mom)-1

plot(net_return.mom, type = "l")
```

Net profit graph
```{r}
plot(net_profit.mom, type='l')
```

Perfomance
```{r}
ann_return.mom <- mean(net_return.mom, na.rm = T)*252*100
std_return.mom <- sd(net_return.mom, na.rm = T)*sqrt(252)*100
sharpe.mom <- ann_return.mom/std_return.mom
sharpe.mom
```


